---
title: "Final Prep"
output: html_notebook
---

# Instructions

- Relax, you got this.
- Complete each task in this Markdown notebook.
- You need not enter anything onto Canvas, because no such thing exists.
- I have included code blocks as needed.
  - Replace `## YOUR CODE HERE!` with your code.
- Some tasks ask you interpretation questions. The goal of these tasks
  is to evaluate your critical thinking skills as they pertain to
  data. You should not need any additional code to complete these
  interpretive tasks.
- Many tasks build upon previous tasks. If you get completely stuck on
  a question, tell me. You won't get any credit for the task I help
  you with, but at least you have a shot at getting credit on the
  tasks which depend on that task.
- This final prep was brought to you by:
    - Metallica's Black Album



# Setup

Run this code chunk to load libraries and import data.

```{r setup, include=FALSE}
library(rpart)
library(rpart.plot)
library(tidyverse)
options(scipen = 999)
appointments <- read_csv("data/appointments.csv")
penguins <- read_csv("data/penguins.csv")
```

# Question Set 1: Penguins

These questions will assess your ability to manipulate data, do some basic plotting and linear regression.

## Data

You all remember our cute bird friends, right?

```{r}
penguins
```

I normally go to great lengths to give you column data, but I don't feel it is necessary here.

## Task 01: Penguin counter!

- Question: How many penguins were on Torgersen island?
- Answer: 


```{r}
## YOUR CODE HERE!

```

## Task 02: Body mass by species

- Question: Which penguin species has the highest average body weight?
- Answer: 

```{r}
## YOUR CODE HERE!

```


## Task 03: Boxplot of body mass by species: 

- Question: Use ggplot to create a box plot of body mass by species.
    - Place `species` on the x axis and `body_mass_g` on the y axis.
- Answer: 
    - Your results should be consistent with your results from Task 02. (Hint)
    - https://www.r-graph-gallery.com/boxplot.html
    - On the final, I'll ask you to upload a plot.
    - You might also want to look up violin plots and understand what they do and how they differ from box plots.

```{r}
## YOUR CODE HERE!

```


## Task 04: Wing length by body mass

Draw a scatter plot using body mass on the x-axis and wing length on the y-axis.

- Question: Do heavier penguins tend to have longer wings?
- Answer: 

```{r}
## YOUR CODE HERE!

```

## Task 05: How many groups?

- Question: If you look at the scatter plot from task 04, you can see that there is a penguin gap right around 205mm of flipper length. This could be because of some bias in our data or sampling error or it could suggest there is more than one group in our dataset. And if that is the case, then there is probably a factor, a variable, which "explains" or "drives" this grouping/separation. Which of the following features (columns) best explains this gap. In other words, which column does the best job of creating a couple of groups? Your options are:
    - sex
    - year
    - species
- Answer: 

```{r}
## YOUR CODE HERE!!!

```


## Task 06: Stratify

Draw a scatter plot using bill_length_mm as the x-axis and bill_depth_mm as the y-axis. Stratify this by species to see if we see similar grouping in bills.

- Question: Do we see two or more fairly distinct groups in this plot as well?
- Answer: 

```{r}
## YOUR CODE HERE!

```

## Task 07: How many gentoos?

- Question: What percentage of birds in our data set are gentoos?
- Answer:

```{r}
## YOUR CODE HERE!

```

And, quite honestly, if all that fails, but you can just get the count of birds by species, there is nothing stopping you from using a calculator to divide 124 by 344. I'm just saying.

## Task 08: Is it a gentoo?

- Question: Create a new column is_gentoo. The values in this column should be 0 if the species is Adlie or Chinstrap and 1 if it is a Gentoo. Confirm that the number of birds identified as TRUE (1) equals the number of Gentoos you found in Task 07.
    - Hello, mutate.
- Answer: 

```{r}
## YOUR CODE HERE!
```


## Task 09: Can you predict a gentoo?

- Question: Using the two models below, which is a better predictor of being a Gentoo? 
    - body mass
    - flipper length
- Answer: 

```{r}
glm_body_mass <- glm(
  is_gentoo~body_mass_g,
  family = "binomial",
  data = penguins
)
summary(glm_body_mass)

glm_flipper_length <- glm(
  is_gentoo~flipper_length_mm,
  family = "binomial",
  data = penguins
)
summary(glm_flipper_length)
```


## Task 10: Accuracy

- Question: Reusing the two models from the previous task, which is a more accurate predictor of being a gentoo and what is that accuracy.
    - I will give you the confusion matrix here AND on the final exam.
    - I don't expect you to come up with this on your own.
- Answer: 

Using the confusion matrices below, your job is to calculate the accuracy of each model.

### Accuracy: Body Mass

```{r}
confusion_data <- penguins %>% select(is_gentoo)
confusion_data$predicted_odds <-
  predict(glm_body_mass, newdata = penguins %>% select(-is_gentoo)) %>% exp()
confusion_data <-
  confusion_data %>%
  mutate(
    predicted_gentoo = case_when(predicted_odds > 1~1,TRUE~0),
    did_not_predict_gentoo = case_when(predicted_odds <= 1~1,TRUE~0),
  )

confusion_data %>%
  group_by(is_gentoo) %>%
  summarize(
    predicted_gentoo = sum(predicted_gentoo),
    did_not_predict_gentoo = sum(did_not_predict_gentoo),
    )
```

### Accuracy: Flipper Length

```{r}
confusion_data <- penguins %>% select(is_gentoo)
confusion_data$predicted_odds <-
  predict(glm_flipper_length, newdata = penguins %>% select(-is_gentoo)) %>% exp()
confusion_data <-
  confusion_data %>%
  mutate(
    predicted_gentoo = case_when(predicted_odds > 1~1,TRUE~0),
    did_not_predict_gentoo = case_when(predicted_odds <= 1~1,TRUE~0),
  )

confusion_data %>%
  group_by(is_gentoo) %>%
  summarize(
    predicted_gentoo = sum(predicted_gentoo),
    did_not_predict_gentoo = sum(did_not_predict_gentoo),
    )
```


## Task 10: Grow a tree

- Question: Using the CaRT model provided below, if a penguin has a flipper length of 208mm, is it (probably) a gentoo?
- Answer: YES!!! 

```{r}
## YOUR CODE HERE!
tree_flipper <- rpart(is_gentoo~flipper_length_mm, data = penguins, method = "class")
rpart.plot(tree_flipper)
```


## Task 11: Just for fun

- Question: Remember our stratified plot (there's that word again) showing THREE groups could be seen using bill length and bill depth. Well, I'd like to show you how flexible CaRT models (decision trees) are. Below is a model using those two features (colulmns).

```{r}
species_bill <- rpart(
  species~bill_length_mm+bill_depth_mm,
  data = penguins,
  method = "class"
)
rpart.plot(species_bill)
```

